Reliable error estimation for Sobol' indices

### 2016-06-17 article.tex
article last changes

\newtheorem{lem}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\theoremstyle{definition}
\newtheorem{algo}{Algorithm}

@@

The problem of interest here is the evaluation of first-order and total effect Sobol' indices. The computation of these indices is performed based on the following integral formula:
\begin{align}
\label{first.order}
\underline{\tau}_u^2  &= \int_{[0,1)^{2d-1}} f(\vx) f(\vx_u:{\vx'}_{-u})d\vx d{\vx'}-\mu^2 , \\
\label{total.effect}
\ov{$\tau$}_u^2 &= \int_{[0,1)^{d+1}}(f(\vx)-f({\vx'}_u:\vx_{-u}))^2d\vx d{\vx'},
\end{align}
where mean and variance of $f$ respectively writes as:
\[ \mu = \int_{[0,1)^{2d}} f(\vx) d{\vx}, \ \qquad
\sigma^2 = \int_{[0,1)^{2d}} f(\vx)^2d{\vx} - \mu^2 .\]
Most of the time, the complexity of $f$ causes the computation of $\mu$, $\sigma^2$ and integrals (\ref{first.order}) and (\ref{total.effect}) to be intractable. In such case, one needs to rely on an estimation of these quantities. In the following, we propose two iterative estimation procedures to evaluate first-order and total-effect Sobol' indices.

@@

We define by design a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1)^d$. Each column of the design refers to a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid [0,1)^{2d}$. One way to estimate the two quantities (\ref{first.order}) and (\ref{total.effect}) is via:
\begin{align}
\label{first.order.est}
\widehat{\underline{\tau}_u^2} & = \frac{1}{n} \sum \limits_{i=0}^{n-1} (f(\vx_i)-\hat{\mu})(f(\vx_{i,u}:{\vx'}_{i,-u})-\hat{\mu}),\\
\label{total.effect.est}
\widehat{\ov{$\tau$}_u^2} & = \frac{1}{2n} \sum \limits_{i=0}^{n-1} (f(\vx_i) - f({\vx'}_{i,u}:\vx_{i,-u}))^2,
\end{align}
using for $\mu$ the efficient estimator introduced in \cite{Monod, Janon}:
\begin{equation}
 \hat{\mu} = \frac{1}{2n} \sum \limits_{i=0}^{n-1} f(\vx_i) + f(\vx_{i,u}:{\vx'}_{i,-u}).
\label{mu.est}
\end{equation}
The estimation of a single pair ($\underline{\tau}_u^2$, $\ov{$\tau$}_u^2$) requires $3n$ evaluations of the model $f$. Using a combinatorial formalism, Saltelli \cite[Theorem 1]{Saltelli} proposed the following strategy to reduce the estimation cost:
\begin{theorem}
\label{saltelli.theorem}
The $d+2$ designs $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i=0}^{n-1}$ constructed for $u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$ allows to estimate all first-order and all total effect Sobol' indices at a cost of $n(d+2)$ evaluations of the model.
\end{theorem}
The $d+2$ designs of Theorem \ref{saltelli.theorem} are obtained by substituting columns of $\mathcal{P}$ for columns of $\mathcal{P}'$ accordingly to subset $u$. While elegant, this approach still requires a number of model evaluations that grows linearly with respect to the input space dimension.

@@

Let $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P}'=\{{\vx'}_i\}_{i=0}^{n-1}$ be two replicated designs of order $1$. Denote by $\{f(\vx_i)\}_{i=0}^{n-1}$ and $\{f({\vx'}_i)\}_{i=0}^{n-1}$ the two sets of model evaluations obtained with $\mathcal{P}$ and $\mathcal{P}'$. Consider $u \subsetneq \mathcal{D}$, from Definition \ref{rep.designs} there exists a permutation $\pi_u$ such that ${\vx'}_{\pi_u(i),u}={\vx}_{i,u}$. Then, remark that $\forall i \in \{0,\dots,n-1\}$:
\[\pi_u(f({\vx'}_i))=f(\vx'_{\pi_u(i),u}:{\vx'}_{\pi_u(i),-u})=f(\vx_{i,u}:{\vx'}_{\pi_u(i),-u})\]
%\pi_u(f({\vx'}_{i,u}:{\vx'}_{i,-u}))
Hence, $\underline{\tau}^2_u$ can be estimated via formula (\ref{first.order.est}) with $\{f(\vx_i)\}_{i=0}^{n-1}$ and $\{\pi_u(f({\vx'}_i))\}_{i=0}^{n-1}$. This estimation method has been further studied and generalized in Tissot et al. \cite{Mara} to the case of closed second-order indices.
\bigskip

@@

Using a recursive estimation procedure requires to write a recursive formula for estimators (\ref{first.order.est}) and (\ref{total.effect.est}). For the sake of readability of the paper, the corresponding recursive formulas are detailed in equations (\ref{first.order.est.rec}) and (\ref{total.effect.est.rec}) in the Annexe.

Procedures $A$ and $B$ also require to iteratively construct two designs $\mathcal{P}_\ell=\{\vx_i\}_{i=0}^{2^\ell-1}$ and $\mathcal{P}'_\ell=\{{\vx'}_i\}_{i=0}^{2^\ell-1}$. These two designs are two Sobol' sequences. Background on Sobol' sequences are provided in \cite{crass}. The construction of $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ is carried out according to the following recursive scheme :
\begin{equation}
\left\lbrace \begin{array}{l}
\mathcal{P}_0= B_0 \\
\mathcal{P}_\ell= \mathcal{P}_{\ell-1} \cup B_\ell \end{array}\right. ,\qquad
\hspace*{0.5cm}
\left\lbrace \begin{array}{l}
\mathcal{P}'_0= {B'}_0 \\
\mathcal{P}'_\ell= \mathcal{P}'_{\ell-1} \cup {B'}_\ell \end{array}\right.,
\label{rec.scheme}
\end{equation}
where at step $\ell\geq0$, $B_\ell$ and $B'_\ell$ are new sets of $2^{\ell-1}$ points added to refine $\mathcal{P}_{\ell-1}$ and $\mathcal{P}'_{\ell-1}$. In the following, $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ refer to the two Sobol' sequences constructed following the multiplicative approach presented in \cite{crass}. This approach guarantee that at each step $\ell\geq0$, $\mathcal{P}_{\ell}$ and $\mathcal{P}'_{\ell}$ are two replicated designs of order $1$.

$\mathcal{P}_{\ell}$ and $\mathcal{P}'_{\ell}$ can either be used with procedure $A$ to estimate all first-order and all total effect Sobol' indices or with procedure $B$ to estimate all first-order Sobol' indices. Algorithms \ref{procedureA} and \ref{procedureB} summarize the main steps of each recursive estimation procedure.
\begin{algorithm}[!ht]
\caption{Procedure $A$}
\begin{algorithmic}[1]
\vspace*{0.2cm}
\State Set: $\ell \leftarrow 1$, $\mathcal{P}_{0} \leftarrow B_0$, ${\mathcal{P}'}_{0} \leftarrow {B'}_0$
\While {stopping criterion}
\State $\mathcal{P}_\ell \leftarrow \mathcal{P}_{\ell-1} \cup B_\ell$

\hspace*{-0.3cm} $\mathcal{P}'_\ell \leftarrow \mathcal{P}'_{\ell-1} \cup {B'}_\ell$
\For {$u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$}
\State Construct design $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i=0}^{2^{\ell}-1}$
\State Compute $\{f(\vx_i)\}_{i=0}^{2^\ell-1}$ and $\{f(\vx_{i,u},{\vx'}_{i,-u})\}_{i=0}^{2^\ell-1}$
\State Estimate $\widehat{\underline{\tau}_u^2}^{(\ell)}$ and $\widehat{\ov{$\tau$}_u^2}^{(\ell)}$ with formula (\ref{first.order.est}) and (\ref{total.effect.est})
\EndFor
\State $\ell \leftarrow \ell + 1$
\EndWhile
\State Return $\widehat{\underline{\tau}_u^2}^{(\ell)}$ and $\widehat{\ov{$\tau$}_u^2}^{(\ell)}$ for all $u \in \mathcal{D}$.
\end{algorithmic}
\label{procedureA}
\end{algorithm}
\begin{algorithm}[!ht]
\caption{Procedure $B$}
\begin{algorithmic}[1]
\vspace*{0.2cm}
\State Set: $\ell \leftarrow 1$, $\mathcal{P}_{0} \leftarrow B_0$, ${\mathcal{P}'}_{0} \leftarrow {B'}_0$
\While {stopping criterion}
\State $\mathcal{P}_\ell \leftarrow \mathcal{P}_{\ell-1} \cup B_\ell$

\hspace*{-0.3cm} $\mathcal{P}'_\ell \leftarrow \mathcal{P}'_{\ell-1} \cup {B'}_\ell$
\State Compute $\{f(\vx_i)\}_{i=0}^{2^\ell-1}$ and $\{f({\vx'}_i)\}_{i=0}^{2^\ell-1}$
\For {$u \in \mathcal{D}$}
\State  $\{f({\vx'}_i)\}_{i=0}^{2^\ell-1}$ into $\{\pi_u(f({\vx'}_i))\}_{i=0}^{2^\ell-1}$
\State Estimate $\widehat{\underline{\tau}_u^2}^{(\ell)}$ with formula (\ref{first.order.est})
\EndFor
\State $\ell \leftarrow \ell + 1$
\EndWhile
\State Return $\widehat{\underline{\tau}_u^2}^{(\ell)}$ for all $u \in \mathcal{D}$.
\end{algorithmic}
\label{procedureB}
\end{algorithm}

The choice of the stopping criterion is the key point of this paper. In most recursive approaches that estimate Sobol' indices, the stopping criterion is a quantity of interest build directly from the estimates. For example, an absolute difference between estimates computed on consecutive steps. Such stopping criteria are often hard to tweak but above all fail to guarantee any error bounds on the estimation. Our stopping criterion is presented and discussed in Section \ref{section.error}. The criterion is an error bound based on the discrete Walsh decomposition of the integral formulas of first-order and total effect Sobol' indices.

@@

The first improvement focus on the use of a better estimator to evaluate small first-order Sobol' indices with procedure $A$. This estimator is called ``Correlation 2" and was introduced by Owen in \cite{Owen}. Owen discussed and highlighted the efficiency of ``Correlation 2" when estimating small closed Sobol' indices.

Our aim is to show that using ``Correlation 2" in procedure $A$ may reduces the total number of model evaluations required while reaching the same accuracy on the estimates.

Cost small: $n(d+1)+d_sn+n$

Cost pas small: $\alpha n d_s +n+(d-d_s)n$

Cost small - cost pas small: $n\sum_{i=1}^{s}(2-\alpha_i+1)$ $s$ number of small indices, $\alpha_i$ proportion of how many more points if not pas small

@@

In the multiplicative approach \cite{crass}, these $2d$ generating matrices are selected as the first $2d$ elements of the set found by Kuo and Joe in \cite{Kuo}. This set is composed of generating matrices that optimize the $2$-dimensional projection $t$-values. For procedure $A$ this choice of these $2d$ generating matrices is logical. Each of the $d+2$ designs constructed following Theorem \ref{saltelli.theorem} is a Sobol' sequence generated by a subset of $d$ matrices among the initial $2d$ ones. Therefore, $2$-dimensional projection $t$-values are not changed.

That is not the case for procedure $B$. (introduire les matrices $C_j$ et ecrire les relations). The permutations applied in step $6$ of Algorithm \ref{procedureB} modify the $2$-dimensional projection $t$-values of the underlying designs.

@@

\section{Applications}
\subsection{Classical test functions}
\subsection{Real case model}

\newpage
\section{Annexe: recursive formulas}

@@

>>>>

\newtheorem{lem}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{algo}{Algorithm}

@@

The problem of interest here is the evaluation of first-order and total effect Sobol' indices. The computation of these indices is performed based on the following integral formula:
\begin{align}
\label{first.order}
\underline{\tau}_u^2  &= \int_{[0,1)^{2d-1}} (f(\vx)-\mu)(f(\vx_{u}:{\vx'}_{-u})-\mu), \\
\label{total.effect}
\ov{$\tau$}_u^2 &= \int_{[0,1)^{d+1}}(f(\vx')-f(\vx_u:{\vx'}_{-u}))^2d\vx d{\vx'}, \qquad u \in \mathcal{D},
\end{align}
where mean and variance of $f$ respectively writes as:
\[ \mu = \int_{[0,1)^{2d}} f(\vx) d{\vx}, \ \qquad
\sigma^2 = \int_{[0,1)^{2d}} f(\vx)^2d{\vx} - \mu^2 .\]
Most of the time, the complexity of $f$ causes the computation of $\mu$, $\sigma^2$ and integrals (\ref{first.order}) and (\ref{total.effect}) to be intractable. In such case, one needs to rely on an estimation of these quantities. In the following, we propose two iterative estimation procedures to evaluate first-order and total-effect Sobol' indices.

@@

We define by design a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1)^d$. Each column of the design refers to a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid [0,1)^{2d}$. One way to estimate the two quantities (\ref{first.order}) and (\ref{total.effect}) is via:
\begin{align}
\label{first.order.est}
\widehat{\underline{\tau}_u^2} & = \frac{1}{n} \sum \limits_{i=0}^{n-1} (f(\vx_i)-\hat{\mu})(f(\vx_{i,u}:{\vx'}_{i,-u})-\hat{\mu}),\\
\label{total.effect.est}
\widehat{\ov{$\tau$}_u^2} & = \frac{1}{2n} \sum \limits_{i=0}^{n-1} (f({\vx'}_i) - f(\vx_{i,u}:{\vx'}_{i,-u}))^2, \qquad u \in \mathcal{D}
\end{align}
using for $\mu$ the efficient estimator introduced in \cite{Monod, Janon}:
\begin{equation}
 \hat{\mu} = \frac{1}{2n} \sum \limits_{i=0}^{n-1} f(\vx_i) + f(\vx_{i,u}:{\vx'}_{i,-u}).
\label{mu.est}
\end{equation}
The estimation of a single pair ($\underline{\tau}_u^2$, $\ov{$\tau$}_u^2$) requires $3n$ evaluations of the model $f$. Using a combinatorial formalism, Saltelli \cite[Theorem 1]{Saltelli} proposed the following strategy to reduce the estimation cost:
\begin{theorem}
\label{saltelli.theorem}
The $d+2$ designs $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i=0}^{n-1}$ constructed for $u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$ allows to estimate all first-order and all total effect Sobol' indices at a cost of $n(d+2)$ evaluations of the model.
\end{theorem}
The $d+2$ designs of Theorem \ref{saltelli.theorem} are obtained by substituting columns of $\mathcal{P}$ for columns of $\mathcal{P}'$ accordingly to $u$. While elegant, this approach still requires a number of model evaluations that grows linearly with respect to the input space dimension.

@@

More specifically, let $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P}'=\{{\vx'}_i\}_{i=0}^{n-1}$ be two replicated designs of order $1$. Denote by $\{f(\vx_i)\}_{i=0}^{n-1}$ and $\{f({\vx'}_i)\}_{i=0}^{n-1}$ the two sets of model evaluations obtained with $\mathcal{P}$ and $\mathcal{P}'$. Consider $u \in \mathcal{D}$, from Definition \ref{rep.designs} there exists a permutation $\pi_u$ such that ${\vx'}_{\pi_u(i),u}={\vx}_{i,u}$. Then, remark that $\forall i \in \{0,\dots,n-1\}$:
\[\pi_u(f({\vx'}_i))=f({\vx'}_{\pi_u(i)})=f(\vx'_{\pi_u(i),u}:{\vx'}_{\pi_u(i),-u})=f(\vx_{i,u}:{\vx'}_{\pi_u(i),-u})\]
%\pi_u(f({\vx'}_{i,u}:{\vx'}_{i,-u}))
Each $\underline{\tau}^2_u$ can be estimated via formula (\ref{first.order.est}) with $\{f(\vx_i)\}_{i=0}^{n-1}$ and $\{\pi_u(f({\vx'}_i))\}_{i=0}^{n-1}$ without requiring further model evaluations. This estimation method has been deeply studied and generalized in Tissot et al. \cite{Mara} to the case of closed second-order indices.
\bigskip

@@

Using a recursive estimation procedure requires to write a recursive formula for estimators (\ref{first.order.est}) and (\ref{total.effect.est}). For the sake of readability of the paper, the corresponding recursive formulas are detailed in equations (\ref{first.order.est.rec}) and (\ref{total.effect.est.rec}) in the Annexe \ref{recursive.formulas}.

Procedures $A$ and $B$ also require to iteratively construct two designs $\mathcal{P}_\ell=\{\vx_i\}_{i=0}^{2^\ell-1}$ and $\mathcal{P}'_\ell=\{{\vx'}_i\}_{i=0}^{2^\ell-1}$. These two designs are two Sobol' sequences. Background on Sobol' sequences are provided in \cite{crass}. We recall here that a $d$-dimensional Sobol' sequence is generated with a set of $d$ generating matrices (one per coordinate). The construction of $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ is carried out according to the following recursive scheme :
\begin{equation}
\left\lbrace \begin{array}{l}
\mathcal{P}_0= B_0 \\
\mathcal{P}_\ell= \mathcal{P}_{\ell-1} \cup B_\ell \end{array}\right. ,\qquad
\hspace*{0.5cm}
\left\lbrace \begin{array}{l}
\mathcal{P}'_0= {B'}_0 \\
\mathcal{P}'_\ell= \mathcal{P}'_{\ell-1} \cup {B'}_\ell \end{array}\right.,
\label{rec.scheme}
\end{equation}
where at step $\ell\geq0$, $B_\ell$ and $B'_\ell$ are new sets of $2^{\ell-1}$ points added to refine $\mathcal{P}_{\ell-1}$ and $\mathcal{P}'_{\ell-1}$. In the following, $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ refer to the two Sobol' sequences constructed following the multiplicative approach presented in \cite{crass}. This approach guarantee that at each step $\ell\geq0$, $\mathcal{P}_{\ell}$ and $\mathcal{P}'_{\ell}$ are two replicated designs of order $1$.

$\mathcal{P}_{\ell}$ and $\mathcal{P}'_{\ell}$ can either be used with procedure $A$ to estimate all first-order and all total effect Sobol' indices or with procedure $B$ to estimate all first-order Sobol' indices. Algorithms \ref{procedureA} and \ref{procedureB} summarize the main steps of each procedure.
\begin{algorithm}[!ht]
\caption{Procedure $A$}
\begin{algorithmic}[1]
\vspace*{0.2cm}
\State Set: $\ell \leftarrow 1$, $\mathcal{P}_{0} \leftarrow B_0$, ${\mathcal{P}'}_{0} \leftarrow {B'}_0$
\While {stopping criterion}
\State $\mathcal{P}_\ell \leftarrow \mathcal{P}_{\ell-1} \cup B_\ell$

\hspace*{-0.3cm} $\mathcal{P}'_\ell \leftarrow \mathcal{P}'_{\ell-1} \cup {B'}_\ell$
\State Compute $\{f(\vx_i)\}_{i=2^{\ell-1}}^{2^\ell-1}$ and $\{f({\vx'}_i)\}_{i=2^{\ell-1}}^{2^\ell-1}$
\For {$u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$}
\State Construct design $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i=2^{\ell-1}}^{2^{\ell}-1}$
\State Compute  $\{f(\vx_{i,u},{\vx'}_{i,-u})\}_{i=2^{\ell-1}}^{2^\ell-1}$
\State Estimate $\widehat{\underline{\tau}_u^2}^{(\ell)}$ and $\widehat{\ov{$\tau$}_u^2}^{(\ell)}$ with formula (\ref{first.order.est.rec}) and (\ref{total.effect.est.rec})
\EndFor
\State $\ell \leftarrow \ell + 1$
\EndWhile
\State Return $\widehat{\underline{\tau}_u^2}^{(\ell)}$ and $\widehat{\ov{$\tau$}_u^2}^{(\ell)}$ for all $u \in \mathcal{D}$.
\end{algorithmic}
\label{procedureA}
\end{algorithm}
\begin{algorithm}[!ht]
\caption{Procedure $B$}
\begin{algorithmic}[1]
\vspace*{0.2cm}
\State Set: $\ell \leftarrow 1$, $\mathcal{P}_{0} \leftarrow B_0$, ${\mathcal{P}'}_{0} \leftarrow {B'}_0$
\While {stopping criterion}
\State $\mathcal{P}_\ell \leftarrow \mathcal{P}_{\ell-1} \cup B_\ell$

\hspace*{-0.3cm} $\mathcal{P}'_\ell \leftarrow \mathcal{P}'_{\ell-1} \cup {B'}_\ell$
\State Compute $\{f(\vx_i)\}_{i=2^{\ell-1}}^{2^\ell-1}$ and $\{f({\vx'}_i)\}_{i=2^{\ell-1}}^{2^\ell-1}$
\For {$u \in \mathcal{D}$}
\State  order $\{f({\vx'}_i)\}_{i=2^{\ell-1}}^{2^\ell-1}$ into $\{\pi_u(f({\vx'}_i))\}_{i=2^{\ell-1}}^{2^\ell-1}$
\State Estimate $\widehat{\underline{\tau}_u^2}^{(\ell)}$ with formula (\ref{first.order.est.rec})
\EndFor
\State $\ell \leftarrow \ell + 1$
\EndWhile
\State Return $\widehat{\underline{\tau}_u^2}^{(\ell)}$ for all $u \in \mathcal{D}$.
\end{algorithmic}
\label{procedureB}
\end{algorithm}

The choice of the stopping criterion is the key point of this paper. In most recursive approaches that estimate Sobol' indices, the stopping criterion is a quantity of interest build directly from the estimates. For example, an absolute difference between estimates computed on consecutive steps. Such stopping criteria are often hard to tweak but above all fail to guarantee any error bound on the estimation. Our stopping criterion is presented and discussed in Section \ref{section.error}. The criterion is an error bound based on the discrete Walsh decomposition of integrals (\ref{first.order}) and (\ref{total.effect}).

@@

The first improvement focus on the use of a better estimator to evaluate small first-order Sobol' indices with procedure $A$. This estimator is called ``Correlation 2" and was introduced by Owen in \cite{Owen}. Owen discussed and highlighted the efficiency of ``Correlation 2" when estimating small closed Sobol' indices. Our aim is to show that using ``Correlation 2" in procedure $A$ may reduces the total number of model evaluations. ``Correlation 2" formula writes as follows:
\begin{equation}
\widehat{\underline{\tau}_u^2} = \frac{1}{n} \sum \limits_{i=0}^{n-1} (f(\vx_i)-f({\vz}_{i,u}:{\vx}_{i,-u}))(f(\vx_{i,u}:{\vx'}_{i,-u})-f({\vx'}_i))
\label{correlation2}
\end{equation}
It uses an extra set of $n$ model evaluations to estimate $\widehat{\underline{\tau}_u^2}$. Owen explains the effectiveness of ``Correlation 2" as follows. If the main effect of $\vx_u$ is small then $f(\vx)$ will be almost entirely explained by $\vx_{-u}$. The two terms $(f(\vx_i)-f({\vz}_{i,u}:{\vx}_{i,-u}))$ and $(f(\vx_{i,u}:{\vx'}_{i,-u})-f({\vx'}_i))$ should then be small quantities as well as (\ref{correlation2}).

We discuss now the potential improvement brought by the use of ``Correlation 2". Assume that the number of small first-order indices is known and equals $\alpha$. Then, the cost of procedure $A$ including ``Correlation 2" to estimate small indices writes: 
\[ n(d-\alpha) + 2 n'\alpha + 2\max(n,n'),\]
where: \begin{itemize}
\item[$\bullet$] $n(d-\alpha)$ is the number of evaluations induced by the $d-\alpha$ designs $\{\vx_{i,u}:{\vx'}_{i,-u}\}_{i\geq0}$ to estimate all non small indices, 
\item[$\bullet$] $2 n'\alpha$ is the number of evaluations induced by the $2\alpha$ designs $\{\vz_{i,u}:\vx_{i,-u}\}_{i\geq0}$ and $\{\vx_{i,u}:{\vx'}_{i,-u}\}_{i\geq0}$ to estimate all small indices,
\item[$\bullet$] $2\max(n,n')$ is the number of evaluations  induced by the $2$ designs $\{\vx_{i}\}_{i\geq0}$ and $\{{\vx'}_{i}\}_{i\geq0}$.
\end{itemize}
At the opposite, the cost of procedure $A$ excluding ``Correlation 2" writes: 
\[ n(d-\alpha) + n\alpha + 2n.\]
It is straightforward that procedure $A$ including ``Correlation 2" performs equally well or better when $n' \leq 2n$. Furthermore, since we are working with Sobol' sequences $n'=2^{k'}$ and $n=2^{k}$. Hence, procedure $A$ including ``Correlation 2" performs equally well or better when the recursive estimation procedure finished at least one step sooner. Illustrations of this improvement will be provided in Section \ref{appli}.

@@

We note by $C_1,\dots,C_d$ the matrices generating $\mathcal{P}_\ell=\{\vx_i\}_{i=0}^{2^\ell-1}$ and ${C'}_1,\dots,{C'}_d$ those generating $\mathcal{P}'_\ell=\{{\vx'}_i\}_{i=0}^{2^\ell-1}$.
In the multiplicative approach \cite{crass}, these $2d$ matrices are selected as the first $2d$ elements of the set found by Kuo and Joe in \cite{Kuo}. This set is composed of generating matrices that optimize the $2$-dimensional projection $t$-values.

%In procedure $A$, each design $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i\geq0}$ constructed in step $6$ of Algorithm \ref{procedureA} is a $d$-dimensional Sobol' sequence generated by the $d$-subset: $${C'}_1,\dots,{C'}_{u-1},C_u,{C'}_{u+1},\dots,C_d.$$ Therefore, $2$-dimensional projection $t$-values are still optimized.

In procedure $B$, each permutation $\pi_u$ applied in step $6$ of Algorithm \ref{procedureB} results in a design $\{\vx_{i,u}:{\vx'}_{\pi_u(i),-u}\}_{i=0}^{2^\ell-1}$. The $d$ generating matrices corresponding to this design are:
$${C'}_1U_u,\dots,{C'}_{u-1}U_u,C_u,{C'}_{u+1}U_u,{C'}_dU_u$$ where $U_u$ is a a full rank upper triangular matrix verifying $C_u={C'}_uU_u$. The right multiplication by $U_u$ is the digital representation of the ordering by $\pi_u$.

@@

\section{Applications}
\label{appli}
\subsection{Classical test functions}
\subsection{Real case model}

\newpage
\section{Annexes}

\subsection{recursive formulas}
\label{recursive.formulas}

@@

\subsection{Proof proposition \ref{right.multi}}
\label{proof.prop}
